---
title: "The Sacred and the Profane: Social, Religious & Medical Effects on Emotional Wellbeing"
author: "Benjamin Villa"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load all the libraries

library(infer)
library(skimr)
library(broom)
library(gganimate)
library(tidyverse)
library(haven)
library(ggplot2)
library(infer)
library(gt)
library(tidyr)
library(readr)
library(purrr)
library(tibble)
library(stringr)
library(forcats)
library(tidyselect)
library(fivethirtyeight)
library(devtools)
library(ggplot2)
library(patchwork)
library(broom)
library(scales)
library(readxl)
library(naniar)
library(janitor)
library(choroplethr)
library(viridis)
library(fuzzyjoin)
library(corrplot)
library(totalcensus)
library(devtools)
library(choroplethrZip)
library(ggfittext)
library(plotly)

```

```{r Data Jobs + Income Read In}

# Read in the data

jobs_raw <- read_excel("Data/jobs_2019.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

income_raw <- read_excel("Data/income.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

```


```{r Gallup Polls Data Read In}

socialwell_index_raw <-read_excel("Data/social_wellbeing_index.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

social_hours_raw <-read_excel("Data/hours_spent_socially.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

religion_important_raw <-read_excel("Data/religion_important_GallupAnalytics.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

church_frequency_raw <- read_excel("Data/church_frequency.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

depression_raw <-read_excel("Data/physician_depression_edit.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

```

```{r Gallup Additional Polls Read In}

emotions_raw <-read_excel("Data/emotions.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

financial_wellbeing_raw <-read_excel("Data/financial_wellbeing.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

alcohol_raw <-read_excel("Data/alcohol.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

party_raw <-read_excel("Data/party.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

religion_raw <-read_excel("Data/religion.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

unemployment_raw <-read_excel("Data/unemployment.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

drugs_med_raw <-read_excel("Data/drugs_med.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

close_relationships_raw <-read_excel("Data/close_relationships.xlsx", sheet = NULL, range = NULL, col_names = TRUE,
  col_types = NULL, na = c("**","*","#"), trim_ws = TRUE, skip = 0,
  n_max = Inf, guess_max = min(1000, 2000),
  progress = readxl_progress(), .name_repair = "unique")

```


```{r Gallup Data Clean}

socialwell_index <- socialwell_index_raw %>%
  clean_names() %>%
  filter(time==2016) %>%
  select(geography,value)
  
social_hours <- social_hours_raw %>%
 clean_names() %>%
  filter(time==2010) %>%
  select(geography,value)

religion_important <- religion_important_raw %>%
 clean_names() %>%
  filter(time==2016) %>%
  select(geography,yes)

church_frequency <- church_frequency_raw %>%
 clean_names() %>%
  filter(time==2016) %>%
  select(geography,at_least_once_a_week)

depression <- depression_raw %>%
  clean_names() %>%
  select(geography,yes,time) %>%
  filter(time==2016) %>%
  rename(depression = yes) %>%
  select(geography,depression)

```

```{r Gallup Data Clean 2}

emotions <- emotions_raw %>%
  clean_names()

financial_wellbeing <- financial_wellbeing_raw %>%
  clean_names()

alcohol <- alcohol_raw %>%
  clean_names()

party <- party_raw %>%
  clean_names()

religion <- religion_raw %>%
  clean_names()

unemployment <- unemployment_raw %>%
  clean_names()

drugs_med <- drugs_med_raw %>%
      select(Geography,`Almost every day Drug Med Mood`,`Sometimes Drug Med Mood`,`Rarely Drug Med Mood`,`Never Drug Med Mood`) %>%
clean_names()

close_relationships <- close_relationships_raw %>%
  clean_names() %>%
  filter(time==2016) %>%
  select(!time)

```

```{r Gallup Join}

# Remove Alcohol?

joined_gallup <- left_join(socialwell_index,social_hours, by="geography") %>%
  left_join(religion_important, by = "geography") %>%
  left_join(church_frequency, by = "geography") %>%
  left_join(depression, by = "geography") %>%
  left_join(emotions, by = "geography") %>%
  left_join(financial_wellbeing, by = "geography") %>%
 # left_join(alcohol, by = "geography") %>%
  left_join(party, by = "geography") %>%
  left_join(religion, by = "geography") %>%
  left_join(unemployment, by = "geography") %>%
  left_join(drugs_med, by = "geography") %>%
  left_join(close_relationships, by = "geography") %>%
  rename(area_title = geography) %>%
  rename(wellbeing_index = value.x) %>%
  rename(social_hours = value.y) %>%
  rename(religion_imp = yes) %>%
  rename(church_week = at_least_once_a_week) %>%
    na.omit()

```

```{r Income Clean}

income <- income_raw %>%
  clean_names() %>%
  rename(area_title = metropolitan_statistical_areas2) %>%
  rename(income_2016 = x2016) %>%
  rename(income_2017 = x2017) %>%
  rename(income_2018 = x2018) %>%
  select(area_title,income_2016)

```

```{r Gallup + Income Joined}

joined_gallup_income <-left_join(joined_gallup, income, by = "area_title")

```


```{r Jobs Clean}

jobs <-jobs_raw %>%
  select(area,area_title,area_type,occ_code,occ_title, tot_emp,jobs_1000_orig,loc_quotient,a_mean)

jobs_religion <- jobs %>%
  filter(str_detect(area_type, "4")) %>%
  filter(str_detect(occ_code, "^21-2")) %>%
  select(area_title,tot_emp,jobs_1000_orig,a_mean) %>%
    na.omit() %>%
  group_by(area_title) %>%
  summarize(tot_emp = sum(tot_emp),
            jobs_1000_orig = sum(as.numeric(jobs_1000_orig)),
            a_mean = sum(a_mean))

jobs_psych <- jobs %>%
  filter(str_detect(area_type, "4")) %>%
  filter(str_detect(occ_code, "^19-303|^29-1223|^31-1133")) %>%
  select(area_title,tot_emp,jobs_1000_orig,a_mean) %>%
    na.omit() %>%
  group_by(area_title) %>%
  summarize(tot_emp = sum(tot_emp),
            jobs_1000_orig = sum(as.numeric(jobs_1000_orig)),
            a_mean = sum(a_mean))

jobs_social <- jobs %>%
    filter(str_detect(area_type, "4")) %>%
  filter(str_detect(occ_code, "^21-10")) %>%
  select(area_title,tot_emp,jobs_1000_orig,a_mean) %>%
    na.omit() %>%
  group_by(area_title) %>%
  summarize(tot_emp = sum(tot_emp),
            jobs_1000_orig = sum(as.numeric(jobs_1000_orig)),
            a_mean = sum(a_mean))

```

```{r Jobs Join}

jobs_joined <- left_join(jobs_social,jobs_psych, by = "area_title", suffix = c("_social","_psych")) %>%
  left_join(jobs_religion, by = "area_title") %>%
  rename(tot_emp_religion = tot_emp) %>%
  rename(jobs_1000_orig_religion = jobs_1000_orig) %>%
  rename(a_mean_religion = a_mean) %>%
    na.omit()

```


```{r Jobs + Gallup + Income Joined}

all_joined <-left_join(joined_gallup_income, jobs_joined, by = "area_title") %>%
   na.omit()


write_rds(all_joined,"shiny/all_joined.rds")

```

```{r Jobs + Gallup + Income Joined}

all_joined2 <-left_join(joined_gallup_income, jobs_joined, by = "area_title") %>%
   na.omit() %>%
   rename(c("Is Religion Important Part of Daily Life?" = "religion_imp",
                           "Attend Church, Synagogue, or Mosque Once Per Week" = "church_week",
                           "Number of Religious Jobs Per 1,000" = "jobs_1000_orig_religion",
                           "Christian" = "christian",
                           "Jewish" = "jewish",
                           "Muslim" = "muslim",
                           
                           "Social Wellbeing Index" = "wellbeing_index",
                           "Hours Spent Socially" = "social_hours",
                           "Very Weak Close Relationship" = "x1_strong_disagree_relationship",
                           "Very Strong Close Relationship" = "x5_strong_agree_relationship",
                           "Number of Social Jobs per 1,000" = "jobs_1000_orig_social",
                           
                           "Republican" = "republican",
                           "Democrat" = "democrat",
                           "Independent / Other" = "independent_other",
                
                           "Income 2016" = "income_2016",
                           "Financial Well-Being" = "value_financial_well_being",
                           "Unemployment Rate" = "value_unemployment",
                           
                           "Happiness" = "yes_happiness",
                           "Enjoyment" = "yes_enjoyment",
                           "Worry" = "yes_worry",
                           "Anger" = "yes_anger",
                           "Sadness" = "yes_sadness", 
                           
                           "Diagnosed Depression" = "depression",
                           "Number of Psych Jobs per 1,000" = "jobs_1000_orig_psych",
                           "Almost Everyday Use Drugs/Medication To Affect Mood" = "almost_every_day_drug_med_mood",
                           "Sometimes Use Drugs/Medication To Affect Mood" = "sometimes_drug_med_mood",
                           "Never Use Drugs/Medication To Affect Mood" = "never_drug_med_mood"))
  


write_rds(all_joined,"shiny/all_joined2.rds")

```



```{r Joined Pivot}

all_pivot <- all_joined %>%
  select(area_title,wellbeing_index,social_hours,religion_imp,church_week,depression,yes_sadness,no_sadness,yes_anger,no_anger,yes_worry,no_worry,yes_enjoyment,no_enjoyment,yes_happiness,no_happiness,value_financial_well_being,
         republican,democrat,independent_other,christian,jewish,muslim,value_unemployment,almost_every_day_drug_med_mood,sometimes_drug_med_mood,rarely_drug_med_mood,never_drug_med_mood, x1_strong_disagree_relationship, x2_disagree_relationship, x3_neutral_relationship, x4_agree_relationship, x5_strong_agree_relationship,
        
       
         income_2016,tot_emp_social,jobs_1000_orig_social,a_mean_social,tot_emp_psych,jobs_1000_orig_psych,a_mean_psych,tot_emp_religion,jobs_1000_orig_religion,a_mean_religion) %>%
  
   pivot_longer(c(wellbeing_index,social_hours,religion_imp,church_week,depression,yes_sadness,no_sadness,yes_anger,no_anger,yes_worry,no_worry,yes_enjoyment,no_enjoyment,yes_happiness,no_happiness,value_financial_well_being,
         republican,democrat,independent_other,christian,jewish,muslim,value_unemployment,almost_every_day_drug_med_mood,sometimes_drug_med_mood,rarely_drug_med_mood,never_drug_med_mood, x1_strong_disagree_relationship, x2_disagree_relationship, x3_neutral_relationship, x4_agree_relationship, x5_strong_agree_relationship,
        
       
         income_2016,tot_emp_social,jobs_1000_orig_social,a_mean_social,tot_emp_psych,jobs_1000_orig_psych,a_mean_psych,tot_emp_religion,jobs_1000_orig_religion,a_mean_religion),
               names_to = "category", values_to= "value"
               )
 
write_rds(all_pivot,"shiny/all_pivot.rds")

```

```{r Any Variable Pivot}

all_pivot %>%
  filter(category == "democrat") %>%
  ggplot(aes(x= reorder(area_title,-value),y = value)) + geom_col(aes(fill = value)) + 
  scale_fill_viridis_c() + 
  xlab(NULL) + ylab(NULL) + theme_minimal() + coord_flip()

```

```{r Any Histogram }

all_pivot %>%
  filter(category == "republican") %>%
  ggplot()+ geom_histogram(aes(value,fill = value)) + 
  scale_fill_viridis_c() + 
  xlab(NULL) + ylab(NULL)


```


```{r Any Correlation}






```



```{r Religious Alone}

all_joined %>%
  ggplot() + geom_histogram(aes(religion_imp), binwidth = 0.05, fill="blue") + theme_classic()
  
all_joined %>%
  ggplot() + geom_histogram(aes(church_week), binwidth = 0.05, fill="blue") + theme_classic()

all_joined %>%
  ggplot(aes(x= reorder(area_title,-church_week),y = church_week)) + geom_col(aes(fill = church_week)) + 
  scale_fill_viridis_c() + 
  xlab(NULL) + ylab(NULL) + theme_minimal() + coord_flip()


```


```{r Correlation Test}

corr <- all_joined %>%
  select(wellbeing_index,social_hours,religion_imp,church_week,depression,yes_sadness,yes_anger,yes_worry,yes_enjoyment,yes_happiness,value_financial_well_being,
         republican,democrat,independent_other,christian,jewish,muslim,value_unemployment,almost_every_day_drug_med_mood,sometimes_drug_med_mood,never_drug_med_mood, x1_strong_disagree_relationship, x5_strong_agree_relationship,
        
       
         income_2016,jobs_1000_orig_social,jobs_1000_orig_psych,jobs_1000_orig_religion) %>%
  rename(c("Is Religion Important Part of Daily Life?" = "religion_imp",
                           "Attend Church, Synagogue, or Mosque Once Per Week" = "church_week",
                           "Number of Religious Jobs Per 1,000" = "jobs_1000_orig_religion",
                           "Christian" = "christian",
                           "Jewish" = "jewish",
                           "Muslim" = "muslim",
                           
                           "Social Wellbeing Index" = "wellbeing_index",
                           "Hours Spent Socially" = "social_hours",
                           "Very Weak Close Relationship" = "x1_strong_disagree_relationship",
                           "Very Strong Close Relationship" = "x5_strong_agree_relationship",
                           "Number of Social Jobs per 1,000" = "jobs_1000_orig_social",
                           
                           "Republican" = "republican",
                           "Democrat" = "democrat",
                           "Independent / Other" = "independent_other",
                
                           "Income 2016" = "income_2016",
                           "Financial Well-Being" = "value_financial_well_being",
                           "Unemployment Rate" = "value_unemployment",
                           
                           "Happiness" = "yes_happiness",
                           "Enjoyment" = "yes_enjoyment",
                           "Worry" = "yes_worry",
                           "Anger" = "yes_anger",
                           "Sadness" = "yes_sadness", 
                           
                           "Diagnosed Depression" = "depression",
                           "Number of Psych Jobs per 1,000" = "jobs_1000_orig_psych",
                           "Almost Everyday Use Drugs/Medication To Affect Mood" = "almost_every_day_drug_med_mood",
                           "Sometimes Use Drugs/Medication To Affect Mood" = "sometimes_drug_med_mood",
                           "Never Use Drugs/Medication To Affect Mood" = "never_drug_med_mood")
    
    
    
    
    
  ) %>%
cor()

write_rds(corr,"shiny/corr.rds")

```


```{r Heat Test Correlation}
corr

corrplot(corr,order = "hclust", tl.offset = 2,tl.cex = .4,
         tl.col = "black")

corrplot(corr, method = "number") # Display the correlation coefficient
corrplot(corr, method = "shade")
corrplot(corr, order = "hclust")

# correlation matrix heat map R 
# http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization

all_joined
  

```


```{r Depression as Dependent}

score_model <- lm(depression ~ wellbeing_index + income_2016 + value_unemployment + almost_every_day_drug_med_mood + `x5_strong_agree_relationship`, data = all_joined) %>%
  tidy(conf.int = TRUE) %>%
  select(term,estimate, conf.low,conf.high)

score_model %>% gt()

```





```{r Religion}

all_joined %>%
  select(religion_imp, church_week, jobs_1000_orig_religion, christian,muslim,jewish,other_dk_refused_religion) %>%
cor() 

```

```{r Negative Emotions}


  select(area_title,yes_sadness,yes_anger,yes_worry,no_enjoyment,no_happiness, no_smile_laugh) 
 
  
emotion_negative <-  all_joined %>%
     select(area_title,yes_sadness,yes_anger,yes_worry,no_enjoyment,no_happiness, no_smile_laugh) %>%
   pivot_longer(c(yes_sadness,yes_anger,yes_worry,no_enjoyment,no_happiness, no_smile_laugh),
               names_to = "emotion", values_to= "value"
               )
 
emotion_negative %>%
  ggplot() + geom_density(aes(value, group=emotion, fill = emotion)) + theme_classic() + facet_wrap(~emotion)

emotion_negative %>%
  filter(area_title == "Akron, OH") %>%
  ggplot() + geom_col(aes(area_title,value, color=emotion), bins = 8, position = position_dodge()) + theme_classic()

```

```{r Positive Emotions}

all_joined %>%
  select(no_sadness,no_anger,no_worry,yes_enjoyment,yes_happiness, yes_smile_laugh) %>%
  cor()

```

```{r}


```


```{r}

```





```{r Graphs Tests}

all_joined %>%
 ggplot(aes(depression,christian)) + geom_point() + geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + theme_classic()

all_joined %>%
  ggplot(aes(depression)) + geom_histogram(binwidth = 0.05, boundary = 0.4, color = "white")

all_joined %>%
rep_sample_n(size=20, reps = 1000)
group_by(replicate) %>%
  summarize(wellbeing_index)

```


```{r Sample Code Regression Write}

dep_psych <- inner_join(jobs_psych, depression, by= "area_title")
write_rds(dep_psych,"shiny/dep_psych.rds")


graph_jobs_income <- income_jobs %>%
ggplot(aes(jobs_1000_orig.x,jobs_1000_orig.y, color=income_2018.x))+ geom_point() + geom_smooth(method = "lm", formula = y ~ x, se = FALSE) +
  scale_y_log10() + scale_x_log10()

score_model <- lm(jobs_1000_orig.x ~ jobs_1000_orig.y + income_2018, data = income_jobs2) %>%
  tidy(conf.int = TRUE)

all_joined %>%
 ggplot(aes(yes_anger,jobs_1000_orig_social)) + geom_point() + geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE)

```


```{r}

```


```{r Notes}

#Map If usable

map<-dict_cbsa %>%
  select(CBSA, CBSA_title,state_full) %>%
  rename(area_title = "CBSA_title")

testing <- all_joined %>%
  left_join(map, by = "area_title") %>%
  rename(region = "CBSA") %>%
  rename(value = "depression") %>%
  distinct()

testing <- distinct(testing, region, .keep_all= TRUE)

zip_choropleth(testing)
data(zip.map)

data("dict_cbsa")


# data i had wasn't able to provide precision needed to provide effect.

```


# About

This webapp is the final project for the GOV 1005. 

## The Sacred and The Profane 

The following project attempts to see if the number of psychiatrists per capita declines in counties where there is an abundant number of clergy per capita. Additionally, the project attempts to look for patterns, correlates, and distributions of clergy and of psychiatrists per capita throughout the United States. 

Research papers include:

“The clergy as a source of mental health assistance: What Americans Believe” by Christopher G. Ellison

“Patterns and Correlates of Contacting Clergy for Mental Disorders in the United States” by Philip S. Wang & Patricia A Berglund

‘Religiousness and mental health: a review’

“The role of religion and spirituality in mental health”

Key Counties to consider include Jefferson County, Alabama, where religious service attendance is 62.9% of the population, Graham County, Arizona, and Ouachita County Arkansas. 

Key links to look into include:

https://www.usatoday.com/story/news/2018/03/13/most-religious-counties-every-state-u-s/421946002/

https://www.jstor.org/stable/20058132?seq=1

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1360908/

https://journals.lww.com/co-psychiatry/Abstract/2014/09000/The_role_of_religion_and_spirituality_in_mental.9.aspx

### Raw Data Sources

The American Medical Association (AMA) Masterfile to calculate the number of psychiatrists, and psychiatric subspecialists practicing in the United States. Additionally, a study "Estimating the Distribution of the US Psychiatric Subespecialist Workforce" published in December 2018.

To calculate the number of Clergy per county, the U.S. Bureau of Labor Statistics was consulted. The data available includes employment of clergy by county, published on May 2018.

Country population has been taken from the United States Census Bureau.

### The map and data explained:
The following maps display the distribution of clergy and psychiatrists per 100,000 citizens in each county in the United States. 

### Walk-Through Video:

Watch the following video to understand the visualization:

#### Created by:
Benjamin Villa Wiesner, a student in the Master in Design Engineering program at Harvard University. 

```{r}

# http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
# https://shiny.rstudio.com/articles/layout-guide.html

# library plotly 

# https://news.gallup.com/poll/161543/provo-orem-utah-religious-metro-area.aspx#2

# https://www.citylab.com/equity/2013/04/americas-most-and-least-religious-metro-areas/5180/

```

